---
title: "Beer Analysis"
author: "Alex Brown"
date: "September 8, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

In this report, I will look into beer production. The data for this report comes form the [Tidy Tuesday Repo](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-03-31) about beer!

## The Data

In this section, I will read in data and do some exploritoy analysis on it.

```{r read-in-data, message=FALSE}

library(tidyverse)

brewing_materials <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewing_materials.csv')
beer_taxed <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_taxed.csv')
brewer_size <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewer_size.csv')
beer_states <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/beer_states.csv')
```

### Brewing Materials

The `brewing_materials` dataset shows the pounds of material used in brewing beer. Here are some observastions I made and steps I took to get the data in the format I want:
* The data has a seperate column for year and month and I will add a column that combines those.


```{r brewing-materials}
brewing_materials <- brewing_materials %>% mutate(year_month = as)
```