---
title: "The Office Analysis"
author: "Alex Brown"
date: "June 17, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This is an analysis of the the office transcripts. The data comes from the [tidytuesday github](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-03-17/readme.md), which points to the `schrute` R package to load data.

## The Data

Below I am loading necissary packages and reading in the data.

```{r read in data, message=FALSE}

# library bank ----

library(data.table)
library(schrute)
library(tidyverse)
library(kableExtra)

# read in data ----

office_transcripts <- theoffice

```

Here is a quick `glimpse` of the data loaded. You can see the `office_transcripts` dataframe represent the transcripts and general information of each episode. In addition, it shows the `imbd` rating and votes of that episode.

```{r about the data}
# about the data ----

glimpse(office_transcripts)

```

When working with a new dataset, it is always a good idea to expose any `null` or `na` values.

```{r find na values}

data.frame(na_values = sapply(office_transcripts, function(x) sum(is.na(office_transcripts)))) %>% rownames_to_column(var = "column_name") %>% kable() %>% kable_styling()

```

I guess that's why they call it "tidy tuesday", there is no missing values!

## Exploring The Data

First, lets get some basic information about the seasons:

### How ma
